---
title: "analysis1"
output: html_document
date: "2024-08-28"
---
TO TRY--detrend data with RW2 model first, then look at residuals vs pm2.5

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(viridis)
library(plotly)
library(scales)
library(pheatmap)
#install.packages("waffle", repos = "https://cinc.rud.is") 
library(reshape2)
#library(HDInterval)
library(rjags)
library(stringr)
library(tidyr)
library(tidybayes)
library(lubridate)

library(tsModel)
library(dlnm)

library(INLA)
library(RColorBrewer)
```

read in data

```{r}

ds0 <- read.csv('./Data/CONFIDENTIAL/AddingFlu_RSV.csv') %>%
  mutate(date =as.Date(date,'%m/%d/%Y'),)

ds1 <- read.csv('./Data/CONFIDENTIAL/CovidNet_AQ_Weather_08292024.csv') %>%
  rename(pm2_5=PM2.5_Avg, humid=site_avg_RelH) %>%
 mutate( date =as.Date(date,'%m/%d/%Y'),
    day_of_week=as.factor(weekdays(date)),
    
    log_Cases_Total=log(Cases_Total+1),
    log_Cases_Total_lag7=lag(log_Cases_Total,7),
    log_Cases_Total_lag14=lag(log_Cases_Total,14),
    
    pm2_5_raw=pm2_5,
          pm2_5=scale(pm2_5_raw),
          pm2_5_lag0=pm2_5,
          pm2_5_lag1 =lag(pm2_5,1),
          pm2_5_lag2 =lag(pm2_5,2),
          pm2_5_lag3 =lag(pm2_5,3),
          pm2_5_lag4 =lag(pm2_5,4),
          pm2_5_lag5 =lag(pm2_5,5),
          pm2_5_lag6 =lag(pm2_5,6),
          pm2_5_lag7 =lag(pm2_5,7),
          pm2_5_lag8 =lag(pm2_5,8),
          pm2_5_lag9 =lag(pm2_5,9),
          pm2_5_lag10 =lag(pm2_5,10),
          pm2_5_lag11 =lag(pm2_5,11),
          pm2_5_lag12 =lag(pm2_5,12),
          pm2_5_lag13 =lag(pm2_5,13),
          pm2_5_lag14 =lag(pm2_5,14),
    
          pm2_5_dic=if_else(pm2_5_raw>15,1,0),
          pm2_5_dic_lag0=pm2_5_dic,
          pm2_5_dic_lag1 = lag(pm2_5_dic,1),
          pm2_5_dic_lag2 = lag(pm2_5_dic,2),
          pm2_5_dic_lag3 = lag(pm2_5_dic,3),
          pm2_5_dic_lag4 = lag(pm2_5_dic,4),
          pm2_5_dic_lag5 = lag(pm2_5_dic,5),
          pm2_5_dic_lag6 = lag(pm2_5_dic,6),
          pm2_5_dic_lag7 = lag(pm2_5_dic,7),
          pm2_5_dic_lag8 = lag(pm2_5_dic,8),
          pm2_5_dic_lag9 = lag(pm2_5_dic,9),
          pm2_5_dic_lag10 = lag(pm2_5_dic,10),
          pm2_5_dic_lag11 = lag(pm2_5_dic,11),
          pm2_5_dic_lag12 = lag(pm2_5_dic,12),
          pm2_5_dic_lag13 = lag(pm2_5_dic,13),
          pm2_5_dic_lag14 = lag(pm2_5_dic,14),

    ave_pm2_5_day_2_9= (pm2_5_lag3 + pm2_5_lag4 + pm2_5_lag5 + pm2_5_lag6 + pm2_5_lag7 + pm2_5_lag8 + pm2_5_lag9)/7,
    
        ave_pm2_5_day_1_5= (pm2_5_lag1 + pm2_5_lag2 + pm2_5_lag3 + pm2_5_lag4 + pm2_5_lag5)/5,
    
    humid_raw=humid,
    humid=scale(humid),
            humid_lag0 =humid,
            humid_lag1=lag(humid,1),
            humid_lag2=lag(humid,2),
            humid_lag3=lag(humid,3),
            humid_lag4=lag(humid,4),
            humid_lag5=lag(humid,5),
            humid_lag6=lag(humid,6),
            humid_lag7=lag(humid,7),
            humid_lag8=lag(humid,8),
            humid_lag9=lag(humid,9),
            humid_lag10=lag(humid,10),
            humid_lag11=lag(humid,11),
            humid_lag12=lag(humid,12),
            humid_lag13=lag(humid,13),
            humid_lag14=lag(humid,14),
             month=as.factor(lubridate::month(date)),
    t=row_number()
          ) %>%
  filter(!is.na(Cases_Total) & !is.na(humid_lag0)) %>%
  left_join(ds0, by='date')

saveRDS(ds1,'./Data/CONFIDENTIAL/lagged_data.rds')

 pm2_5_mat <- ds1 %>%
   select(starts_with('pm2_5_lag')) %>%
   as.matrix()
 
  pm2_5_mat_dic <- ds1 %>%
   select(starts_with('pm2_5_dic_lag')) %>%
   as.matrix()
 
  humid_mat <- ds1 %>%
   select(starts_with('humid_lag')) %>%
 
 
 dow_mat <- model.matrix(~-1 + day_of_week, data=ds1) 
 dow_mat <- dow_mat[,-1] #remove intercept

  month_mat <- model.matrix(~-1 + month, data=ds1) 
  month_mat <- month_mat[,-1]

  
```

```{r}
plot(ds1$ave_pm2_5_day_2_9,ds1$Cases_Total)
```

```{r}
mod1 <- mgcv::gam(Cases_Total~ day_of_week + month +   s(t) + s(ave_pm2_5_day_1_5, bs='ps'), family='poisson', data=ds1)

 plot(mod1)
 abline(h=0)
 abline(v=20)

mod1 <- mgcv::gam(Cases_Total~  s(t, bs='ps') +   ave_pm2_5_day_1_5 +day_of_week, family='poisson', data=ds1)

#library("splines")

ds1$pm2_5_round=round(ds1$ave_pm2_5_day_1_5,1)

mod2 <- inla(Cases_Total~ f(t, model='rw2', scale.model=T) + ave_pm2_5_day_1_5 + day_of_week , family='nbinomial', data=ds1)

 mod2 <- inla(Cases_Total~ f(t, model='rw2', scale.model=T) + pm2_5_lag1 + pm2_5_lag2 +pm2_5_lag3 +pm2_5_lag4 +pm2_5_lag5 + day_of_week , family='nbinomial', data=ds1)

coefs <- mod2$summary.fixed
coefs.lag <- coefs[grep('lag',row.names(coefs)),'mean']

plot(coefs.lag, type='l')
abline(h=0)

plot(mod2)

# plot(mod2$summary.random$t$mean, ds1$pm2_5_raw) #doesn't seem to be correlation between PM2.5 and time rnadom effect
# abline(v=0)

# mod1 <- mgcv::gam(RSV_Total~ day_of_week + month +  s(pm2_5_raw, bs='ps'), family='poisson', data=ds1)
# plot(mod1)
```



## Distributed lag linear models
whether you adjust time using a monthly dummy or using a RW2 makes a huge effect on estimates of PM2.5 effect
```{r}
  nlag=14

  basis_pm2_5_lin <-   crossbasis(ds1$pm2_5, lag=nlag, argvar=list(fun="lin"),
      arglag=list(fun="poly",degree=4))
  
  colnames(basis_pm2_5_lin) = paste0("basis_pm2_5.", colnames(basis_pm2_5_lin))


  ds1a <- ds1 %>%
    mutate(weekdate=floor_date(date, 'week'),
           weekN = as.numeric(difftime(weekdate, min(weekdate), units = "weeks")),
           cal_monthN= as.numeric(as.factor(paste(year(date), month(date), '01', sep='-')))
           )
  
mod1_lin <-   inla(Cases_Total ~ day_of_week  + basis_pm2_5_lin + f(t, model='rw2', scale.model=T), family='nbinomial', data=ds1a,
                      control.predictor = list(link = 1, compute = TRUE),
                   control.compute=list(return.marginals.predictor=TRUE),
                control.fixed = list(correlation.matrix = TRUE, 
                            prec.intercept = 1e-4, prec = 1e-4)
               )
summary(mod1_lin)

mod1.coef <- mod1_lin$summary.fixed[grep('basis_pm',row.names(mod1_lin$summary.fixed)),'mean']

mod1.vcov <-mod1_lin$misc$lincomb.derived.covariance.matrix[grep('basis_pm',row.names(mod1_lin$summary.fixed)),grep('basis_pm',row.names(mod1_lin$summary.fixed))]

pred1.pm <- crosspred(basis_pm2_5_lin, coef=mod1.coef, vcov=mod1.vcov, at=-10:40,bylag=0.2, cumul=TRUE)


plot(pred1.pm, "overall", var=1,lag=1, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
main="log-RR")

plot(pred1.pm, "slices", var=1, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
main="Association with an 1 SD increase in PM2.5")

 plot(pred1.pm, "slices", var=1, col=2, cumul=TRUE, ylab="Cumulative RR",
main="Cumulative association with a 1-SD increase in PM2.5")

```

DLNM model
DLNM https://github.com/gasparrini/hydromet_dengue/blob/main/04_lag_nonlinear_output.R

```{r}
  nlag=14
  pm2_5_lags <- Lag(ds1$pm2_5,k=0:nlag)
  
  basis_pm2_5 <- crossbasis(pm2_5_lags,
                    argvar = list(fun = "ns", knots = equalknots(ds1$pm2_5, 2)),
                    arglag = list(fun = "ns", knots = nlag/2))
  
  colnames(basis_pm2_5) = paste0("basis_pm2_5.", colnames(basis_pm2_5))

  humid_lags <- Lag(ds1$humid,k=0:nlag)
  
    basis_humid <- crossbasis(humid_lags,
                    argvar = list(fun = "ns", knots = equalknots(ds1$humid, 2)),
                    arglag = list(fun = "ns", knots = nlag/2))
  colnames(basis_humid) = paste0("basis_humid", colnames(basis_humid))
  
mod1 <-   inla(Cases_Total ~ day_of_week + month + basis_pm2_5 + basis_humid, family='poisson', data=ds1,
                      control.predictor = list(link = 1, compute = TRUE),
                control.fixed = list(correlation.matrix = TRUE, 
                            prec.intercept = 1e-4, prec = 1e-4)
               )
# extract full coef and vcov and create indicators for each term
coef <- mod1$summary.fixed$mean
vcov <- mod1$misc$lincomb.derived.covariance.matrix

# find position of the terms associated with Tmin crossbasis
indt <- grep("basis_pm2_5", mod1$names.fixed)

# extract predictions from the Tmin DLNM centred on overall mean Tmin (19 deg C)

predt <- crosspred(basis_pm2_5, coef = coef[indt], vcov=vcov[indt,indt],
                    bylag = 0.25, cen = round(mean(ds1$pm2_5), 2), cumul=T) 


plot(predt, "overall", var=1, col=3, ylab="logRR", ci.arg=list(density=15,lwd=2),
main="log-RR")

plot(predt, "slices", var=1, lag=2,col=3, ylab="logRR", ci.arg=list(density=15,lwd=2),
main="Association with an 1 SD increase in PM2.5")

 plot(predt, "slices", var=1,lag=3, col=2, cumul=TRUE, ylab="Cumulative log_RR",
main="Cumulative association with a 1-SD increase in PM2.5")

y <- predt$predvar
x <- seq(0, nlag, 0.25)
z <- t(predt$matRRfit)

pal <- rev(brewer.pal(11, "PRGn"))
levels <- pretty(z, 20)
col1 <- colorRampPalette(pal[1:6])
col2 <- colorRampPalette(pal[6:11])
cols <- c(col1(sum(levels < 1)), col2(sum(levels > 1)))

filled.contour(x,y,z,
               xlab = "Lag", ylab = 'Humidity', main = "",
               col = cols,levels = levels,
               plot.axes = { axis(1, at = 0:nlag, c(0:nlag)) 
                 axis(2)})
mtext(side = 2, at = max(y)*1.1, text = "a", las = 2, cex = 1.2, line = 2)
```

hhh4 model
```{r}
library(surveillance)

cases <- ds1 %>% 
    reshape2::dcast(date~., value.var= 'Cases_Total') %>%
  dplyr::select(-date) %>%
    as.matrix() 

pm2_5_mat <- ds1 %>% 
    reshape2::dcast(date~., value.var= 'pm2_5_lag5') %>%
  dplyr::select(-date) %>%
    as.matrix() 

mod_str1 <- list(
      end = list(f = ~ 1 + day1 +day2 + day3 +day4+day5+day6 +pm2_5_mat ),
      ar = list(f = ~ 1  ),
      ne = list(f = ~ -1 ), #single location, no neighbords
      family = "NegBin1",
      data=list(pm2_5_mat=pm2_5_mat,
                day1=dow_mat[,1,drop=F],
                day2=dow_mat[,2,drop=F],
                day3=dow_mat[,3,drop=F],
                day4=dow_mat[,4,drop=F],
                day5=dow_mat[,5,drop=F],
                day6=dow_mat[,6,drop=F])
    )

 #Define STS object
first.date=min(ds1$date)
start.month=lubridate::month(first.date)
start.year=lubridate::year(first.date)

  covid_df <- sts(cases, start = c(start.year, start.month), frequency = 365)
  
  hhh4_fit <- hhh4(stsObj = covid_df, control = mod_str1)
  
  summary(hhh4_fit)

```


## try to detrend first, then look at association
```{r}
mod1 <-   inla(Cases_Total ~ day_of_week + f(t, model='rw2', scale.model=F) , family='poisson', data=ds1,
                      control.predictor = list(link = 1, compute = TRUE),
                control.fixed = list(correlation.matrix = TRUE, 
                            prec.intercept = 1e-4, prec = 1e-4)
               )

plot(mod1)

preds <- exp(mod1$summary.linear.predictor$mean)


ds1 <- ds1 %>%
  mutate(detrended = Cases_Total / preds)

ggplot(ds1) +
  geom_point(x=detrended, y=pm2_5_lag2)

plot(ds1$pm2_5_lag5,log(ds1$detrended))
abline(h=0)
cor(ds1$pm2_5_lag5,(ds1$detrended))

cor(ds1$pm2_5_lag5,preds)

```



### Define the model 
```{r}

model_string <- "

model{
for(i in 1:Ntimes){

  cases[i] ~ dnegbin(prob[i],r)
  prob[i]<- r/(r+lambda[i])  ## likelihood 

  log(lambda[i]) <-   alpha[i] +inprod(delta[],dow_mat[i,]) +
                    inprod(epsilon[],pm2_5_mat[i,]) +
                    inprod(eta[],   humid_mat[i,]) +
                    inprod(theta[],   pm2_5_mat[i,]) 
                    
  

}

#AR1 random intercept
  alpha[1] ~ dnorm(mu0  , (1 - rho1^2) * tau.alpha1) # Prior for the first time point
  for(i in 2:Ntimes){
    alpha[i] ~ dnorm( mu0  + rho1 * (alpha[i-1]-mu0), tau.alpha1)
  }
  
  #day of week effects
  for(k in 1:6){
    delta[k]~dnorm(0,1e-4)
  }
  
  
  epsilon[1] ~ dnorm(0, (1 - rho2^2) * tau.epsilon2)
  for(m in 2:15){
    epsilon[m] ~ dnorm(rho2 *epsilon[m-1] , tau.epsilon2)
  }
  
  eta[1] ~ dnorm(0, (1 - rho3^2) * tau.eta2)
  for(m in 2:15){
    eta[m] ~ dnorm(rho3 *eta[m-1] , tau.eta2)
  }
  
    theta[1] ~ dnorm(0, (1 - rho4^2) * tau.theta2)
  for(m in 2:15){
    theta[m] ~ dnorm(rho4 *theta[m-1] , tau.theta2)
  }
  

  tau.alpha1 ~ dgamma(3, 2)  # Tight prior for tau 
  tau.epsilon2 ~ dgamma(3, 2)  # Tight prior for tau 
  tau.eta2 ~ dgamma(3, 2)  # Tight prior for tau 
  tau.theta2 ~ dgamma(3, 2)  # Tight prior for tau 

  rho1 ~ dunif(0, 1)       # Uniform prior for rho
  rho2 ~ dunif(0, 1)       # Uniform prior for rho
  rho3 ~ dunif(0, 1)       # Uniform prior for rho
  rho4 ~ dunif(0, 1)       # Uniform prior for rho

  mu0 ~ dnorm(0, 1e-4)   # Uninformative prior for intercept
  
  r ~ dunif(0,250)

}
"


##SIMPLIFIED VERSION

model_string_simp <- "

model{
for(i in 1:Ntimes){

  cases[i] ~ dnegbin(prob[i],r)
  prob[i]<- r/(r+lambda[i])  ## likelihood 

  log(lambda[i]) <-    alpha + epsilon[i] +
                    inprod(delta[], dow_mat[i,]) +
                    inprod(theta[],   pm2_5_mat[i,]) 
}

   #day of week effects
  for(k in 1:6){
    delta[k]~dnorm(0,1e-4)
  }
  
  
  epsilon[1] <- 0
  epsilon[2] ~ dnorm(epsilon[1], tau.epsilon2)  # Second value depends on first

      for(m in 3:Ntimes){
        epsilon[m] ~ dnorm(2 * epsilon[m-1] - epsilon[m-2], tau.epsilon2)
      }
  
  alpha~ dnorm(0,1e-4)
  
 theta[1] ~ dnorm(0, (1 - rho2^2) * tau.theta2)
  for(m in 2:15){
    theta[m] ~ dnorm(rho2 *theta[m-1] , tau.theta2)
  }

  tau.theta2 ~ dgamma(3, 2)  # Tight prior for tau 

  #tau.epsilon <- exp(log.tau.epsilon)
  tau.epsilon2 ~ dgamma(1, 5e-5) #this is what INLA uses

  rho1 ~ dunif(0, 1)       # Uniform prior for rho
  rho2 ~ dunif(0, 1)       # Uniform prior for rho

  mu0 ~ dnorm(0, 1e-4)   # Uninformative prior for intercept
  
  r ~ dunif(0,250)

}
"
```

### Set random seeds. 
We are going to run 3 chains in the MCMC, so we need a seed for each. This ensures reproducible results.

```{r}
inits1=list(".RNG.seed"=c(123), ".RNG.name"='base::Wichmann-Hill')
inits2=list(".RNG.seed"=c(456), ".RNG.name"='base::Wichmann-Hill')
inits3=list(".RNG.seed"=c(789), ".RNG.name"='base::Wichmann-Hill')
```

### Initialize the model for COVID
```{r}
model_spec<-textConnection(model_string_simp)
model_jags<-jags.model(model_spec,
                       inits=list(inits1,inits2, inits3),
                       data=list('Ntimes'= nrow(ds1)  ,
                                 'cases'= ds1$Cases_Total,
                                 'dow_mat'=dow_mat,
                                 'pm2_5_mat'=pm2_5_mat_dic
                       ),
                       n.adapt=15000,
                       n.chains=3, quiet=F)
```

### Posterior sampling

what parameters do you want to sample?
```{r}
params<-c('rho1', 'mu0','lambda', 'epsilon', 'theta')

```

```{r}
posterior_samples<-coda.samples(model_jags, 
                                params, 
                                n.iter=10000)

saveRDS(posterior_samples,'./Data/covid_posteriors.rds')

```

## examine lambda
```{r}
test <- gather_draws(posterior_samples, lambda[i])

test %>%
  filter(i==1) %>%
  mutate(.chain=as.factor(.chain)) %>%
ggplot( aes(x=.iteration, y=.value, group=.chain, color=.chain), alpha=0.2) +
  geom_line()+
  theme_classic()


```

```{r}
test <- gather_draws(posterior_samples, epsilon[i])

test %>%
  filter(i==1) %>%
  mutate(.chain=as.factor(.chain)) %>%
ggplot( aes(x=.iteration, y=.value, group=.chain, color=.chain), alpha=0.2) +
  geom_line()+
  theme_classic()


```



HDI for lambda
```{r}
lambda_quant <- test %>%
  median_hdi()

ggplot(lambda_quant, aes(x=i, y=.value)) +
  geom_line()+
  theme_minimal()
```

distributed lag theta (pm2.5)
```{r}
dl_summary <- gather_draws(posterior_samples, theta[m]) %>%
  median_hdi()

ggplot(dl_summary, aes(x=m, y=.value)) +
  geom_line()+
  geom_ribbon(aes(x=m, ymin=.lower, ymax=.upper), alpha=0.2)+
  theme_classic()+
  geom_hline(yintercept=0, lty=2)


humid_draws_covid <- gather_draws(posterior_samples, theta[m]) 

humid_cum_effect_covid <- humid_draws_covid %>%
  arrange(.chain, .iteration, m) %>%
  group_by(.chain, .iteration) %>%
  mutate(cum_effect= cumsum(.value)) %>%
  ungroup() %>%
  dplyr::select(m,.chain, .iteration, cum_effect)%>%
  group_by(m) %>%
  median_hdi()

ggplot(humid_cum_effect_covid, aes(x=m, y=cum_effect)) +
  geom_line()+
  geom_ribbon(aes(x=m, ymin=.lower, ymax=.upper), alpha=0.2)+
  theme_classic()+
  geom_hline(yintercept=0, lty=2)



```
RW2 component

```{r}
rw_draws_covid <- gather_draws(posterior_samples, epsilon[i]) 

rw_effect_covid <- rw_draws_covid %>%
  arrange(.chain, .iteration, i) %>%
  ungroup() %>%
  dplyr::select(i,.chain, .iteration, .value)%>%
  group_by(i) %>%
  median_hdi()

ggplot(rw_effect_covid, aes(x=i, y=.value)) +
  geom_line()+
  geom_ribbon(aes(x=i, ymin=.lower, ymax=.upper), alpha=0.2)+
  theme_classic()+
  geom_hline(yintercept=0, lty=2)
```



## Same model for Flu

### Initialize the model for FLU
```{r}
model_spec<-textConnection(model_string_simp)
model_jags_flu<-jags.model(model_spec,
                       inits=list(inits1,inits2, inits3),
                       data=list('Ntimes'= nrow(ds1)  ,
                                 'cases'= ds1$Flu_Total,
                                 'month_mat'=month_mat,
                                 'dow_mat'=dow_mat,
                                 'humid_mat'=humid_mat,
                                 'pm2_5_mat'=pm2_5_mat
                       ),
                       n.adapt=15000,
                       n.chains=3, quiet=F)
```

### Posterior sampling

what parameters do you want to sample?
```{r}
params<-c('rho1', 'mu0','alpha','tau.alpha1','lambda', 'epsilon','eta', 'theta')

```

```{r}
posterior_samples_flu<-coda.samples(model_jags_flu, 
                                params, 
                                n.iter=10000)

saveRDS(posterior_samples_flu,'./Data/flu_posteriors.rds')
```

## examine lambda
```{r}
test_flu <- gather_draws(posterior_samples_flu, lambda[i])

test_flu %>%
  filter(i==1) %>%
  mutate(.chain=as.factor(.chain)) %>%
ggplot( aes(x=.iteration, y=.value, group=.chain, color=.chain), alpha=0.2) +
  geom_line()+
  theme_classic()


```

```{r}
test_flu2 <- gather_draws(posterior_samples_flu, epsilon[i])

test_flu2 %>%
  filter(i==1) %>%
  mutate(.chain=as.factor(.chain)) %>%
ggplot( aes(x=.iteration, y=.value, group=.chain, color=.chain), alpha=0.2) +
  geom_line()+
  theme_classic()


```

```{r}
test_flu3 <- gather_draws(posterior_samples_flu, eta[i])

test_flu3 %>%
  filter(i==1) %>%
  mutate(.chain=as.factor(.chain)) %>%
ggplot( aes(x=.iteration, y=.value, group=.chain, color=.chain), alpha=0.2) +
  geom_line()+
  theme_classic()


```

HDI for lambda
```{r}
lambda_quant <- test_flu3 %>%
  median_hdi()

ggplot(lambda_quant, aes(x=i, y=.value)) +
  geom_line()+
  theme_minimal()
```


### distributed lag for FLU
```{r}
dl_summary_flu <- gather_draws(posterior_samples_flu, epsilon[m]) %>%
  median_hdi()

ggplot(dl_summary_flu, aes(x=m, y=.value)) +
  geom_line()+
  geom_ribbon(aes(x=m, ymin=.lower, ymax=.upper), alpha=0.2)+
  theme_minimal()+
  geom_hline(yintercept=0, lty=2)

test_flu <- gather_draws(posterior_samples_flu, epsilon[i])

test_flu %>%
  filter(i==1) %>%
  mutate(.chain=as.factor(.chain)) %>%
ggplot( aes(x=.iteration, y=.value, group=.chain, color=.chain), alpha=0.2) +
  geom_line()+
  theme_classic()
```

distributed lag eta (humidity)
```{r}
dl_summary_flu <- gather_draws(posterior_samples_flu, eta[m]) %>%
  median_hdi()

ggplot(dl_summary_flu, aes(x=m, y=.value)) +
  geom_line()+
  geom_ribbon(aes(x=m, ymin=.lower, ymax=.upper), alpha=0.2)+
  theme_classic()+
  geom_hline(yintercept=0, lty=2)

humid_draws_flu <- gather_draws(posterior_samples_flu, eta[m]) 

humid_cum_effect_flu <- humid_draws_flu%>%
  arrange(.chain, .iteration, m) %>%
  group_by(.chain, .iteration) %>%
  mutate(cum_effect= cumsum(.value)) %>%
  ungroup() %>%
  dplyr::select(m,.chain, .iteration, cum_effect)%>%
  group_by(m) %>%
  median_hdi()

ggplot(humid_cum_effect_flu, aes(x=m, y=cum_effect)) +
  geom_line()+
  geom_ribbon(aes(x=m, ymin=.lower, ymax=.upper), alpha=0.2)+
  theme_classic()+
  geom_hline(yintercept=0, lty=2)
```

distributed lag eta (PM2.5)
```{r}
dl_summary_flu <- gather_draws(posterior_samples_flu, theta[m]) %>%
  median_hdi()

ggplot(dl_summary_flu, aes(x=m, y=.value)) +
  geom_line()+
  geom_ribbon(aes(x=m, ymin=.lower, ymax=.upper), alpha=0.2)+
  theme_classic()+
  geom_hline(yintercept=0, lty=2)

```


### AR1 component
```{r}
ar1_summary_flu <- gather_draws(posterior_samples_flu, alpha[i]) %>%
  median_hdi()

ggplot(ar1_summary_flu, aes(x=i, y=.value)) +
  geom_line()+
  geom_ribbon(aes(x=i, ymin=.lower, ymax=.upper), alpha=0.2)+
  theme_minimal()

```



## Same model for RSV

### Initialize the model for COVID
```{r, eval=F}
model_spec<-textConnection(model_string_simp)
model_jags_RSV<-jags.model(model_spec,
                       inits=list(inits1,inits2, inits3),
                       data=list('Ntimes'= nrow(ds1)  ,
                                 'cases'= ds1$RSV_Total,
                                 'month_mat'=month_mat,
                                 'dow_mat'=dow_mat,
                                 'humid_mat'=humid_mat,
                                 'pm2_5_mat'=pm2_5_mat
                       ),
                       n.adapt=15000,
                       n.chains=3, quiet=F)
```

### Posterior sampling

what parameters do you want to sample?
```{r}
params<-c('rho1', 'mu0','alpha','tau.alpha1','lambda', 'epsilon','eta', 'theta')

```

```{r, eval=F}
posterior_samples_RSV<-coda.samples(model_jags_RSV,
                                    params,
                                    n.iter=10000)
saveRDS(posterior_samples_RSV,'./Data/rsv_posteriors.rds')

```
## examine lambda
```{r}
test_RSV <- gather_draws(posterior_samples_RSV, lambda[i])

test_RSV %>%
  filter(i==1) %>%
  mutate(.chain=as.factor(.chain)) %>%
  ggplot( aes(x=.iteration, y=.value, group=.chain, color=.chain), alpha=0.2) +
  geom_line()+
  theme_classic()


```

```{r}
test_RSV2 <- gather_draws(posterior_samples_RSV, epsilon[i])

test_RSV2 %>%
  filter(i==1) %>%
  mutate(.chain=as.factor(.chain)) %>%
  ggplot( aes(x=.iteration, y=.value, group=.chain, color=.chain), alpha=0.2) +
  geom_line()+
  theme_classic()


```

```{r}
test_RSV3 <- gather_draws(posterior_samples_RSV, eta[i])

test_RSV3 %>%
  filter(i==1) %>%
  mutate(.chain=as.factor(.chain)) %>%
  ggplot( aes(x=.iteration, y=.value, group=.chain, color=.chain), alpha=0.2) +
  geom_line()+
  theme_classic()


```

HDI for lambda
```{r}
lambda_quant <- test_RSV3 %>%
  median_hdi()

ggplot(lambda_quant, aes(x=i, y=.value)) +
  geom_line()+
  theme_minimal()
```

### distributed lag for RSV
```{r}
dl_summary_RSV <- gather_draws(posterior_samples_RSV, epsilon[m]) %>%
  median_hdi()

ggplot(dl_summary_RSV, aes(x=m, y=.value)) +
  geom_line()+
  geom_ribbon(aes(x=m, ymin=.lower, ymax=.upper), alpha=0.2)+
  theme_minimal()+
  geom_hline(yintercept=0, lty=2)

test_RSV <- gather_draws(posterior_samples_RSV, epsilon[i])

test_RSV %>%
  filter(i==1) %>%
  mutate(.chain=as.factor(.chain)) %>%
ggplot( aes(x=.iteration, y=.value, group=.chain, color=.chain), alpha=0.2) +
  geom_line()+
  theme_classic()
```

distributed lag eta (humidity)
```{r}
dl_summary_RSV <- gather_draws(posterior_samples_RSV, eta[m]) %>%
  median_hdi()

ggplot(dl_summary_RSV, aes(x=m, y=.value)) +
  geom_line()+
  geom_ribbon(aes(x=m, ymin=.lower, ymax=.upper), alpha=0.2)+
  theme_classic()+
  geom_hline(yintercept=0, lty=2)

```


### AR1 component
```{r}
ar1_summary_RSV <- gather_draws(posterior_samples_RSV, alpha[i]) %>%
  median_hdi()

ggplot(ar1_summary_RSV, aes(x=i, y=.value)) +
  geom_line()+
  geom_ribbon(aes(x=i, ymin=.lower, ymax=.upper), alpha=0.2)+
  theme_minimal()
```
simple..
```{r}

ds2 <- ds1 %>%
  arrange(date) %>%
  mutate(flu_lag1=lag(Flu_Total,1),
         log_flu_lag1 =log(flu_lag1+1),
         monthN=as.factor(lubridate::month(date))) %>%
  dplyr::select(date, Cases_Total,starts_with('humid'),starts_with('pm2_5_lag0'),)
mod1 <- list()

humid.select <- c( 'humid_lag1','humid_lag2','humid_lag3','humid_lag4','humid_lag5','humid_lag6','humid_lag7','humid_lag8','humid_lag9',
                  'humid_lag10','humid_lag11','humid_lag12','humid_lag13')

pm2_5.select <- c( 'pm2_5_lag0','pm2_5_lag1','pm2_5_lag2','pm2_5_lag3','pm2_5_lag4','pm2_5_lag5','pm2_5_lag6','pm2_5_lag7','pm2_5_lag8','pm2_5_lag9',  'pm2_5_lag10','pm2_5_lag11','pm2_5_lag12','pm2_5_lag13')

pm2_5.select2 <- c( 'pm2_5_dic','pm2_5_dic_lag1','pm2_5_dic_lag2','pm2_5_dic_lag3','pm2_5_dic_lag4','pm2_5_dic_lag5','pm2_5_dic_lag6','pm2_5_dic_lag7','pm2_5_dic_lag8','pm2_5_dic_lag9',  'pm2_5_dic_lag10')

mod1 <- list()

for(i in 1:13){
  ds2$humidvar <- ds2[,humid.select[i]]
mod1[[i]] <- MASS::glm.nb(Cases_Total ~ humidvar +monthN, data=ds2)
}

humid_coefs <- sapply(mod1, function(X){
  coef(X)['humidvar']
    })

plot(1:13,humid_coefs, type='l')
abline(h=0)


#pm2.5
mod1 <- list()

for(i in 1:11){
  ds2$pm25_var <- ds2[,pm2_5.select2[i]]
mod1[[i]] <- MASS::glm.nb(Cases_Total ~ pm25_var +monthN + day_of_week, data=ds2)
}

pm2_5_coefs <- sapply(mod1, function(X){
  coef(X)['pm25_var']
    })

plot(1:11,pm2_5_coefs, type='l')
abline(h=0)

mod2 <- MASS::glm.nb(Cases_Total ~ pm2_5_dic_lag4 + pm2_5_dic_lag5 + pm2_5_dic_lag6 +monthN +day_of_week, data=ds2)
summary(mod2)
```

```{r}
ds2 <- ds1 %>%
  mutate(std_humid = (humid - mean(humid) )/sd(humid)) %>%
  arrange(region, date) %>%
  group_by(date) %>%
  mutate(  
           std_humid_lag0 = std_humid,
           std_humid_lag1 = lag(std_humid,1),
           std_humid_lag2 = lag(std_humid,2),
           )
```

