---
title: "josh summary"
format: pdf
editor: visual
---

```{r, setup}
library(dplyr)
library(ggplot2)
library(scales)
library(reshape2)
library(rjags)
library(stringr)
library(tidyr)
library(tidybayes)
library(lubridate)

library(tsModel)
library(dlnm)

library(INLA)
library(RColorBrewer)

ds1 <- readRDS('./Data/CONFIDENTIAL/lagged_data.rds')
```

## Create basis functions

```{r}
nlag=14

basis_pm2_5_lin <-   crossbasis(ds1$pm2_5, lag=nlag, argvar=list(fun="lin"),
                                arglag=list(fun="poly",degree=4))

colnames(basis_pm2_5_lin) = paste0("basis_pm2_5.", colnames(basis_pm2_5_lin))


ds1a <- ds1 %>%
  mutate(weekdate=floor_date(date, 'week'),
         weekN = as.numeric(difftime(weekdate, min(weekdate), units = "weeks")),
         cal_monthN= as.numeric(as.factor(paste(year(date), month(date), '01', sep='-')))
  )
```

## INLA model 1

Day of week effect

PM2.5 distributed lag basis

RW2 using time index

Negative binomial

```{r}

mod1_lin <-   inla(Cases_Total ~ day_of_week  + basis_pm2_5_lin + f(t, model='rw2', scale.model=T), family='nbinomial', data=ds1a,
                   control.predictor = list(link = 1, compute = TRUE),
                   control.compute=list(return.marginals.predictor=TRUE),
                   control.fixed = list(correlation.matrix = TRUE, 
                                        prec.intercept = 1e-4, prec = 1e-4)
)
```

Extract coefficients

```{r}
mod1.coef <- mod1_lin$summary.fixed[grep('basis_pm',row.names(mod1_lin$summary.fixed)),'mean']

mod1.vcov <-mod1_lin$misc$lincomb.derived.covariance.matrix[grep('basis_pm',row.names(mod1_lin$summary.fixed)),grep('basis_pm',row.names(mod1_lin$summary.fixed))]

pred1.pm <- crosspred(basis_pm2_5_lin, coef=mod1.coef, vcov=mod1.vcov, at=-10:40,bylag=0.2, cumul=TRUE, model.link='log')

```

## Plot 

Trends

```{r}
plot(pred1.pm, "overall", var=1,lag=1, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
     main="log-RR")
```

Coefficients

```{r}
plot(pred1.pm, "slices", var=1, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
     main="Association with an 1 SD increase in PM2.5")
```

Cumulative effect
```{r}
plot(pred1.pm, "slices", var=1, col=2, cumul=TRUE, ylab="Cumulative RR",
     main="Cumulative association with a 1-SD increase in PM2.5")
```


## INLA model 2

Same as INLA model 1, except we have a RW2 for month index (instead of day inde)

Day of week effect

PM2.5 distributed lag basis

RW2 using time index

Negative binomial

```{r}

mod2_lin <-   inla(Cases_Total ~ day_of_week  + basis_pm2_5_lin + f(cal_monthN, model='rw2', scale.model=T), family='nbinomial', data=ds1a,
                   control.predictor = list(link = 1, compute = TRUE),
                   control.compute=list(return.marginals.predictor=TRUE),
                   control.fixed = list(correlation.matrix = TRUE, 
                                        prec.intercept = 1e-4, prec = 1e-4)
)
```

Extract coefficients

```{r}
mod2.coef <- mod2_lin$summary.fixed[grep('basis_pm',row.names(mod1_lin$summary.fixed)),'mean']

mod2.vcov <-mod2_lin$misc$lincomb.derived.covariance.matrix[grep('basis_pm',row.names(mod2_lin$summary.fixed)),grep('basis_pm',row.names(mod2_lin$summary.fixed))]

pred2.pm <- crosspred(basis_pm2_5_lin, coef=mod2.coef, vcov=mod2.vcov, at=-10:40,bylag=0.2, cumul=TRUE, model.link='log')

```

## Plot 

Trends

```{r}
plot(pred2.pm, "overall", var=1,lag=1, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
     main="log-RR")
```

Coefficients

```{r}
plot(pred2.pm, "slices", var=1, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
     main="Association with an 1 SD increase in PM2.5")
```

Cumulative effect
```{r}
plot(pred2.pm, "slices", var=1, col=2, cumul=TRUE, ylab="Cumulative RR",
     main="Cumulative association with a 1-SD increase in PM2.5")
```


## INLA model 3
Same as model 1 but use dichotomous PM2.5

### Create basis functions

```{r}
nlag=14

basis_pm2_5_dic_lin <-   crossbasis(ds1$pm2_5_dic, lag=nlag, argvar=list(fun="lin"),
                                arglag=list(fun="poly",degree=4))

colnames(basis_pm2_5_dic_lin) = paste0("basis_pm2_5.", colnames(basis_pm2_5_dic_lin))


ds1a <- ds1 %>%
  mutate(weekdate=floor_date(date, 'week'),
         weekN = as.numeric(difftime(weekdate, min(weekdate), units = "weeks")),
         cal_monthN= as.numeric(as.factor(paste(year(date), month(date), '01', sep='-')))
  )
```


Day of week effect

PM2.5 distributed lag basis

RW2 using time index

Negative binomial

```{r}

mod3_lin <-   inla(Cases_Total ~ day_of_week  + basis_pm2_5_dic_lin + f(t, model='rw2', scale.model=T), family='nbinomial', data=ds1a,
                   control.predictor = list(link = 1, compute = TRUE),
                   control.compute=list(return.marginals.predictor=TRUE),
                   control.fixed = list(correlation.matrix = TRUE, 
                                        prec.intercept = 1e-4, prec = 1e-4)
)
```

Extract coefficients

```{r}
mod3.coef <- mod3_lin$summary.fixed[grep('basis_pm',row.names(mod3_lin$summary.fixed)),'mean']

mod3.vcov <-mod3_lin$misc$lincomb.derived.covariance.matrix[grep('basis_pm',row.names(mod3_lin$summary.fixed)),grep('basis_pm',row.names(mod3_lin$summary.fixed))]

pred3.pm <- crosspred(basis_pm2_5_dic_lin, coef=mod3.coef, vcov=mod3.vcov, at=-10:40,bylag=0.2, cumul=TRUE, model.link='log')

```

## Plot 

Trends

```{r}
plot(pred3.pm, "overall", var=1,lag=1, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
     main="log-RR")
```

Coefficients

```{r}
plot(pred3.pm, "slices", var=1, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
     main="Association with an 1 SD increase in PM2.5")
```

Cumulative effect
```{r}
plot(pred3.pm, "slices", var=1, col=2, cumul=TRUE, ylab="Cumulative RR",
     main="Cumulative association with a 1-SD increase in PM2.5")
```



##Model 4

Same as mod1 but use GAM instead of INLA

```{r}
nlag=14

basis_pm2_5_lin <-   crossbasis(ds1$pm2_5, lag=nlag, argvar=list(fun="lin"),
                                arglag=list(fun="poly",degree=4))

colnames(basis_pm2_5_lin) = paste0("basis_pm2_5.", colnames(basis_pm2_5_lin))


ds1a <- ds1 %>%
  mutate(weekdate=floor_date(date, 'week'),
         weekN = as.numeric(difftime(weekdate, min(weekdate), units = "weeks")),
         cal_monthN= as.numeric(as.factor(paste(year(date), month(date), '01', sep='-')))
  )
mod4_lin <-   mgcv::gam(Cases_Total ~ day_of_week  + basis_pm2_5_lin + s(t, bs='ps'), family='nb', data=ds1a)

```

Extract coefficients

```{r}

pred4.pm <- crosspred(basis_pm2_5_lin, model=mod4_lin, cumul=TRUE)

```

## Plot 

Trends

```{r}
plot(pred4.pm, "overall", var=1,lag=1, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
     main="log-RR")
```

Coefficients

```{r}
plot(pred4.pm, "slices", var=1, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
     main="Association with an 1 SD increase in PM2.5")
```

Cumulative effect
```{r}
plot(pred4.pm, "slices", var=1, col=2, cumul=TRUE, ylab="Cumulative RR",
     main="Cumulative association with a 1-SD increase in PM2.5")
```
